---
author: Salif Nikiema
title: "ProjectPart1"
output: html_document
---

```{r}
library(tidyverse)
library(dplyr)
```

```{r}
cfb <- read_csv("messy_dataset.csv")
```

#Introduction 
For my Data set I selected a college football team statistics i shortened the data from its original file because the original contained 150 different variables which is understanding because football contains many stats but I managed to trim it down to about 6 core variables and I intentionally added some missing variables.
-Team
-Games
-Offensive Yards
-Passing Yards
-Rush Yards
-Points Per Game

#Research Questions 
1.Do teams with higher offensive yards tend to win more games
2.Is passing yards or rushing yards more strongly related to points per game
3.Do teams with higher overall offensive production score more points per game?

#DataSet
```{r}
cfb
```
#Checking for missing values and duplicates 
```{r}
# checking for missing values by column
colSums(is.na(cfb))

# checking for duplicates
sum(duplicated(cfb))
```

#Cleaning

```{r}
# Remove duplicates if I have any 
cfb_clean <- cfb %>%
  distinct() %>%
  mutate(`Off Yards`= ifelse(is.na(`Off Yards`),median(`Off Yards`, na.rm = TRUE), `Off Yards`),
    `Pass Yards` = ifelse(is.na(`Pass Yards`),median(`Pass Yards`,na.rm = TRUE), `Pass Yards`),
    `Rush Yds` = ifelse(is.na(`Rush Yds`), median(`Rush Yds`, na.rm = TRUE), `Rush Yds`),
    `Points Per Game` = ifelse(is.na(`Points Per Game`), median(`Points Per Game`, na.rm = TRUE), `Points Per Game`),
    Games  = ifelse(is.na(Games), median(Games, na.rm = TRUE), Games))


```

#Double Checking to see if it works
```{r}
colSums(is.na(cfb_clean))
sum(duplicated(cfb_clean))
```
#Making A csv of my clean dataset
```{r}
readr::write_csv(cfb_clean, "cleaned_dataset.csv")
```


#Codebook 

- *Team*:  Team name (categorical)  
- *Games*: Number of games played (numeric)  
- *Off Yards*:Total offensive yards (numeric)  
- *Pass Yards*: Total passing yards (numeric)  
- *Rush Yds*:  Total rushing yards (numeric)  
- *Points Per Game*: Average points per game (numeric)  



#Documentation 

1.I cut down the original data set from 150 plus variables to 6 key variables. 

2.I added some missing values to my data set and added a duplicate row.

3.When cleaning i used the distinct function to remove duplicates and replaced numeric values the column median.

#Principles of DM

1. Data quality - I checked my dataset for missing values and duplicates and I cleaned it by removing duplicates and placing in the missing values

2.Data Governance - I kept track of all the changes that i had made to the original data set by including the missing data and removing a bunch of variables 

#Principles of FM

1.Specific naming - I have my files named  so that way i know which files are which.

2.Organized- all my files for my project are organized in the same folder.

#LLM Usage 

I used LLM in order to help me clean my data and verify that the code worked 
